Red Wine Quality Analyst By Mohamed Hassona 
========================================================



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.
knitr::opts_chunk$set( echo=FALSE,warning=FALSE,message=FALSE)
library(ggplot2)
library(dplyr)
library(knitr)
# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:/Users/teaca/Documents/r project")
pf <- read.csv("wineQualityReds.csv")
pf$X <- NULL
```



# Univariate Plots Section



```{r echo=FALSE, Univariate_Plots1}
str(pf)
```

```{r echo=FALSE, Univariate_Plots2}
summary(pf)
```

The dataset has 12 variables and nearly 1600 observations  


```{r echo=FALSE, Univariate_Plots3}
# Histogram of quality
ggplot(aes(x = quality), data = pf) +
  geom_histogram(binwidth=1)
```

The majority of wine quality is between 5 and 6 this is a normal distrubution this shows that the majority of wines are of average quility with very few scoring exceptionly high or low 



```{r echo=FALSE, Univariate_Plots4}
# Histogram of alcohol
ggplot(aes(x = alcohol), data = pf) +
  geom_histogram(binwidth=0.5)
```

The mean and median make this a normal disrtribution but the graph looks like it is right skewed 

```{r echo=FALSE, Univariate_Plots5}
# Histogram of sulphates
ggplot(aes(x = sulphates), data = pf) +
  geom_histogram(binwidth=0.1)
```

Sulphates like alchol has a mean and median make this a normal disrtribution but the graph looks like it is right skewed

```{r echo=FALSE, Univariate_Plots6}
# Histogram of pH
ggplot(aes(x = pH), data = pf) +
  geom_histogram(binwidth=0.1)
```

```{r echo=FALSE, Univariate_Plots7}
# boxplot of pH
ggplot(aes(y = pH), data = pf) +
  geom_boxplot()
```

The pH values have a normal distribution with very few outliers 


```{r echo=FALSE, Univariate_Plots8}
# Histogram of citric acid
ggplot(aes(x = citric.acid), data = pf) +
  geom_histogram(binwidth=0.05)
```

```{r echo=FALSE, Univariate_Plots9}
# Histogram of citric acid without outliers
ggplot(aes(x = citric.acid), data = pf) +
  geom_histogram(binwidth=0.05)+
  scale_x_continuous(lim = c(0, 0.75), breaks = seq(0, 0.75, 0.05))
  
```

Citric acid is right skewed I made a secound histrogram to look closer at the lower values and ignor the higher outliers

```{r echo=FALSE, Univariate_Plots10}
# Histogram of volatile.acidity
ggplot(aes(x = volatile.acidity), data = pf) +
  geom_histogram(binwidth=0.05)
```
```{r echo=FALSE, Univariate_Plots12}
# Histogram of volatile.acidity without outliers
ggplot(aes(x = volatile.acidity), data = pf) +
  geom_histogram(binwidth=0.05)+
  scale_x_continuous(lim = c(0, 1.2), breaks = seq(0, 1.2, 0.5))
```

The mean and median make this a normal disrtribution the first graph looks like it is right skewed but I took out outliers for the secound graph and you can see it's a normal distrobution 

```{r echo=FALSE, Univariate_Plots13}
# Histogram of fixed.acidity
ggplot(aes(x = fixed.acidity), data = pf) +
  geom_histogram(binwidth=0.5)
```

Fixed acidity is slightly right skewed and most values are between 7-8 

```{r echo=FALSE, Univariate_Plots14}
# Histogram of chlorides
ggplot(aes(x = chlorides), data = pf) +
  geom_histogram(binwidth=0.01)
```

This is intresting this graph has a high standard devation and is right skewed and most values are close to 0.08 and a max value of 0.6


# Univariate Analysis



### What is the structure of your dataset?
The dataset has 12 variables and 1599 observations
### What is/are the main feature(s) of interest in your dataset?
The main variable is quality I would like to see how it is determined by other variables 
### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?
I think alchol will correlate well with quality 
### Did you create any new variables from existing variables in the dataset?
I did not make any new variables 
### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
alchol and sulphates had an unusual distribution which had a right skew and normal mean/median (alchol and sulphates), on some of the graphs I reduced the x axcise to take out outliers and to loook at the data in more detail  

# Bivariate Plots Section





```{r echo=FALSE, Bivariate_Plots1}
# scatter of alcohol - quality with jitter 
ggplot(aes(x = alcohol, y = quality), data = pf) +
  geom_jitter(alpha=0.5, width = 0.1, height = 0.35)
```


```{r echo=FALSE, Bivariate_Plots2}
cor.test(pf$quality, pf$alcohol)
```

I added jitter to this graph to make it more clear for the relationship between quality and alcohol the correlation coefficient is 0.476 this shows that there is a stong relationship, the graph shows that higher alcohol content genrally have higher quality rateing 

```{r echo=FALSE, Bivariate_Plots3}
cor.test(pf$pH, pf$quality)
```


```{r echo=FALSE, Bivariate_Plots4}
# scatter of pH - quality
ggplot(aes(x = pH, y = quality), data = pf) +
  geom_jitter(alpha=0.5, width = 0, height = 0.35)
```

I added jitter to this grah to make it more clear, pH and quality have a low negative correlation coefficient of -0.0577 this shows they have almost no relationship, looking at the graph allot of wines with the highest quality rateing have a pH value between 3 - 3.25


```{r echo=FALSE, Bivariate_Plots5}
cor.test(pf$fixed.acidity, pf$density)
```

```{r echo=FALSE, Bivariate_Plots6}
# scatter of fixed.acidity - density
ggplot(aes(x = fixed.acidity, y = density), data = pf) +
  geom_point()
```

Fixed acidity and density have a correlation coefficient of 0.668 this shows a very stong relationship, the graph shows a clear trend between thease 2 variables 



# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
Investigation. How did the feature(s) of interest vary with other features in \
the dataset?
I looked at a number of relationships alchol and quility showed a promising relationship 
### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
I found density and fixed acidity had a storng relationship 
### What was the strongest relationship you found?
Density and fixed acidity had the stongest relationship with a correlation coefficient of 0.668

# Multivariate Plots Section



```{r echo=FALSE, Multivariate_Plots1}
# scatter of fixed.acidity - density by quality
ggplot(aes(x = fixed.acidity, y = density), data = pf) +
  geom_point(aes(color = as.factor(quality)))+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)
```

I thought it might be intresting to see quality added to this graph as density and fixed acidity had a high corrilation, from this graph it looks like density and fixed acidity do not have much effect on quality 

```{r echo=FALSE, Multivariate_Plots2}
# scatter of  alcohol - sulphates by quality
ggplot(aes(x = alcohol, y = sulphates), data = pf) +
  geom_jitter(alpha=0.5, width = 0.1, height = 0, aes(color = as.factor(quality)))+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)
```

This graph shows somthing intresting allot of the high quility wine is between 0.75 - 1 sulphates 

```{r echo=FALSE, Multivariate_Plots3}
# scatter of volatile.acidity - quality by alcohol
ggplot(aes(x = volatile.acidity, y = quality), data = pf) +
  geom_jitter(alpha=0.5, width = 0, height = 0.5, aes(color = alcohol))

```

This graph shows that most wine with a quility score or 8 have a volatile acidity of less than 0.5

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
Investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
There are a few things that high rated wine have the stongest relationship that I found was alcohol content with higher being better 
### Were there any interesting or surprising interactions between features?
Quality rateing 8 wine with a volatile acidity 0.5 and lower seemed to have the lowest alcohol content in that rateing


# Final Plots and Summary



### Plot One
```{r echo=FALSE, Plot_One}
# hitrogram of quality
ggplot(data=pf, aes(factor(quality))) + 
  geom_bar(colour="black", fill = "steelblue") +
  ggtitle("Quality of Wine Histrogram")+
  xlab("Quality") + 
  ylab("Number Of Wines")
```

### Description One
The majorty of wine is rated 5 or 6 with very few rated higher or lower it is rated on a scale of 0 - 10 with no entrys for 0-2 and 9-10, this is a normal distrubution this shows that the majority of wines are of average quility with very few scoring exceptionly high or low 

### Plot Two
```{r echo=FALSE, Plot_Two}
# scatter of fixed.acidity - density 
ggplot(aes(x = fixed.acidity, y = density), data = pf) +
  geom_point(colour="steelblue")+
  ggtitle("Density and Fixed Acidity Scatter Graph")+
  xlab(expression(Fixed~Acidity~(g/dm^{3}))) + 
  ylab(expression(Density~(g/cm^{3})))


```

### Description Two
Fixed acidity and density have a correlation coefficient of 0.668 this shows a very stong relationship, the graph shows a clear trend between thease 2 variables 


### Plot Three
```{r echo=FALSE, Plot_Three}
#scatter of alcohol - sulphates by quality
ggplot(aes(x = alcohol, y = sulphates), data = pf) +
  geom_jitter(alpha=0.5, width = 0.3, height = 0, aes(color = as.factor(quality)))+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
  ggtitle("Alcohol and Sulphates Scatter Graph by Quality")+
  xlab("Alcohol (% by Volume)") + 
  ylab(expression(Sulphates~(g/dm^{3})))

```

### Description Three

This shows the correlation quality has with higher alcohol and suphate range between 0.75 and 1

------
# Observations 
This data set was intresting to look though it has 12 variables related to wine. going though I had some insights into what makes a high quality wine the main being alcohol, haveing a sulphates between 0.75 and 1 also corraltes with high rateings. I also found that density and fixed acidity corrlated stongly. 

# Reflection
At first I found R to be hard to get into but I have found it to b a very good programming launguage for makeing graphs I still prefer jupter notebooks to r studio. this project has been very good at putting all the code I have learnt into practice and I have learnt allot about putting a project toghther  
  
